---
title: "Covid91 vaccine study Final2024"
author: "Estefania Maldonado"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))

male = filter(FinalData, sex =="M")
female = filter(FinalData, sex =="F")
lgbtq = filter(FinalData, LGBTQ =="gay")
drugUser = filter(FinalData, DrugUser =="yes")
```


# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Methods
Looking at the data collected, this will be categorical explained by categorical statistical problem, we will create separate data variables for each group to separate and test each of them. We will use the candP helper file to use R studios built in bar charts to display data graphically. W will also use xtabs to find the percentages relevant to the data. In this case we are only comparing two types of data and therefore we do not need to use any loops to measure multiple variables in the Fisher test or run a FisherXY.

## Males

### Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment,data=male)
barchartGC(~infected + treatment,data=male, type="percent")
```
  
The bar chart above shows the graphical outcomes for males, based on the collected data. It compares the rates of contracting Covid-19 as compared to vaccination status. The left shows people who contracted Covid-19 and the right shows men who have not.

This clearly shows that more than twice the percentage of people who are in the placebo group contracted Covid where people who did not contract Covid are very closely related.

### Numerical Descriptive Results

```{r}
tableMale <- xtabs(~infected + treatment, data=male)
rowPerc(tableMale)
colPerc(tableMale)
```

The set of data illustrates a twofold difference in individuals who received the vaccinations and encountered Covid-19. The placebo group contracted Covid approximately 74% of the time where the vaccinated group contracted Covid 26% of the time. However, there was no significant contrast among those who received vaccinations and were deemed "OK." We can see that the vaccine is generally effective in this testing scenario.


### Inferential Results

```{r}
chisq.test(tableMale)
chisqtestGC(tableMale)
maleFisher <- fisher.test(tableMale)
maleFisher
```

   The Chi Squared test shows that the p-value for males is 0.0000000000001098. This is substantially smaller than the general use point of 0.05. Consequently, the very small number indicates that there is a difference in the data. The Fisher's Exact test yielded a p-value of 0.00000000000004631, signifying a significant difference in the occurrence of Covid-19 in the data. This all shows that the odds a unvaxinated man getting Covid are 2.2 times higher than a man with the vaccine contracting Covid.

### Confidence Interval

```{r}
oddsMale = maleFisher$estimate/(maleFisher$estimate+1)*100
print(oddsMale)
maleLow=(maleFisher$conf.int[1]/(maleFisher$conf.int[1]+1))*100
maleHigh=(maleFisher$conf.int[2]/(maleFisher$conf.int[2]+1))*100
cat("\n","Condifence interval for Males: ",maleLow," to ",maleHigh)
```

The confidence interval is >50% with a normal vaccine threshold of >60%. This confidence interval is more than adequate to show that this is a good vaccine for this group.

## Females


## LGBTQ



## Druggies


# Overall Results and Conclusions